Machine Learning guided workflow for Ribosome Binding Site engineering

### 2021-05-17 paper/main.tex
Updated paper to new story.

The \textbf{Story} (3 sentences):
\begin{enumerate}
    \item we automate the design of sequences for finding high TIR RBS
    \begin{enumerate}
        \item DBTL
    \end{enumerate}
    \item the ML method discovers biological knowledge (validates existing knowledge), including the highest TIR ever
    \begin{enumerate}
        \item high G
        \item diverse library
        \item edit distance $>$ 2
    \end{enumerate}
    \item BioFoundry helps ML (good data), and ML helps BioFoundry (more automation)
    \begin{enumerate}
        \item Sequences have small technical variation
        \item UCB much more efficient than random
    \end{enumerate}
\end{enumerate}

@@

We show our DBTL workflow in Figure \ref{fig: Flowchart}.
% with the machine learning process emphasised.
BUILD and TEST are driven by choices made by human researchers.
Machine learning algorithms are applied in two parts.
In the DESIGN phase, we designed experimental design policy to recommend RBS sequences in batches so that we can identify RBS sequences with highest possible TIR. 
% We use the Upper Confidence Bound multi-armed Bandit algorithm to recommend sequences to be Built and Tested next.
In LEARN phase, we use the Gaussian Process regression algorithm to predict the TIR of different RBS sequences.
The goal is to provide predictions about TIR labels and uncertainty to assist the recommendation.

@@

\subsection{Performance of the recommendation algorithm}
In our genetic design, the investigated RBS controls expression of the Green Fluorescent Protein (GFP) in its mut3b variant. 
By controlling expression of a fluorescent protein with the RBS we can quickly assess the perceived relative TIR by measuring fluorescence of cells harbouring plasmid with the device over time.
Finally, the mRNA is transcribed from an IPTG-inducible promoter pLlacO-1. 
By making the whole device inducible we can synchronise the start of the expression of the GFP in all the cultures by inducing them at the same time with addition of IPTG.

@@

The design recommendations were made using the Multi-armed Bandit algorithm.
In short, the this algorithm is a stochastic method of probing of the experimental space. 
This algorithm aims at maximising the reward (output) from testing a limited number of instances from a big pool which cannot be wholly tested due to limited resources (time, computational power, capital). 
In our case we use the Upper Confidence Bound algorithm, which focuses its recommendations on sequences that should give highest TIR based on the probabilities computed by the prediction algorithm (GPR). 
Another feature of the bandits algorithms is that it balances two parts: the exploration in terms of potential high TIRs in unknown areas, and exploitation in terms of querying areas which is known to give relatively high TIRs.
% The two approaches can be controlled with the \textbeta\enspace parameter. 
% We have decided that in the first iterations of the cycle it would be beneficial to skew the algorithm towards the exploration with exploitation taking increasing role in later iterations. 
One thing of note is that the bandit algorithm is stochastic, that is it exploits the probabilities of given event occurring (in this case RBS having a specific TIR). 
As such, it pairs naturally with our prediction algorithm, the Gaussian Process, which provides probability based function regression.

@@

\subsection{Automated RBS variant construction}

To help us obtain reliable and reproducible results we have employed automation-heavy workflow for our experiments.
This way we were able to eliminate a big part of sample-to-sample variation as well as human-introduced variation.
Additionally, performing all the procedures directly in 96-well microplate format enabled us to significantly cut down time to prepare our variants.\\
In short, the genetic variations of the RBS were introduced to the plasmids with combination of PCR and isothermal assembly. 
The plasmids were then transformed and the resulting transformants were tested using microplate reader.
Vast majority of reactions were prepared using liquid handling equipment.
Similarly, colony picking was done by an automated colony picker.\\
Figure \ref{fig: SD} shows the distribution of relative error for each tested RBS, calculated as the average TIR from 6 biological replicates divided by the given samples standard deviation (coefficient of variation).
\mengyan{Maciej: can you explain what is "given samples standard deviation"?}
The phenomenon of higher volatility of readings in the lower end of fluorescence measurements is well documented and comes from higher influence of small changes on the percentage error in that range.
Overall, 79\% and 27\% of our samples shows relative error of below 40\% and 20\% respectively which we find satisfactory for the purposes of this study.

@@

>>>>

@@

\subsection{The experimental workflow}

We show our DBTL workflow in Figure \ref{fig: Flowchart}.
% with the machine learning process emphasised.
BUILD and TEST are driven chiefly by choices made by human researchers and use of automated methods.
Machine learning algorithms are applied in DESIGN AND LEARN.
In the DESIGN phase, we designed experimental design policy to recommend RBS sequences in batches so that we can identify RBS sequences with highest possible TIR. 
% We use the Upper Confidence Bound multi-armed Bandit algorithm to recommend sequences to be Built and Tested next.
In LEARN phase, we use the Gaussian Process regression algorithm to predict the TIR of RBS sequences comprising the experimental space.
The goal is to provide predictions about TIR labels and uncertainty to assist the recommendation.\\

@@

To help us obtain reliable and reproducible results we have employed automation-heavy workflow for our experiments.
This way we were able to eliminate a big part of sample-to-sample variation as well as human-introduced variation.
Additionally, performing all the procedures directly in 96-well microplate format enabled us to significantly cut down time to prepare our variants.\\
In short, the genetic variations of the RBS were introduced to the plasmids with combination of PCR and isothermal assembly. 
The plasmids were then transformed and the resulting transformants were tested using microplate reader.
Vast majority of reactions were prepared using liquid handling equipment.
Similarly, colony picking was done by an automated colony picker.\\

\subsection{Design of the investigated genetic device}

In our genetic design, the investigated RBS controls expression of the Green Fluorescent Protein (GFP) in its mut3b variant. 
By controlling expression of a fluorescent protein with the RBS we can quickly assess the perceived relative TIR by measuring fluorescence of cells harbouring plasmid with the device over time.
Finally, the mRNA is transcribed from an IPTG-inducible promoter pLlacO-1. 
By making the whole device inducible we can synchronise the start of the expression of the GFP in all the cultures by inducing them at the same time with addition of IPTG.

@@

\subsection{Performance of the recommendation algorithm}

The design recommendations were made using the Multi-armed Bandit algorithm.
In short, this algorithm is a stochastic method of probing of the experimental space. 
This algorithm aims at maximising the reward (output) from testing a limited number of instances from a big pool which cannot be wholly tested due to limited resources (time, computational power, capital). 
In our case we use the Upper Confidence Bound algorithm, which focuses its recommendations on sequences that should give highest TIR based on the probabilities computed by the prediction algorithm (GPR). 
Another feature of the bandits algorithms is that it balances two parts: the exploration in terms of potential high TIRs in unknown areas, and exploitation in terms of querying areas which is known to give relatively high TIRs.
% The two approaches can be controlled with the \textbeta\enspace parameter. 
% We have decided that in the first iterations of the cycle it would be beneficial to skew the algorithm towards the exploration with exploitation taking increasing role in later iterations. 
One thing of note is that the bandit algorithm is stochastic, that is it exploits the probabilities of given event occurring (in this case RBS having a specific TIR). 
As such, it pairs naturally with our prediction algorithm, the Gaussian Process, which provides probability based function regression.

@@

@@

\subsection{Advantages of pairing Machine Learning with lab automation}

Figure \ref{fig: SD} shows the distribution of relative error for each tested RBS, calculated as the average TIR from 6 biological replicates divided by the given samples standard deviation (coefficient of variation).
\mengyan{Maciej: can you explain what is "given samples standard deviation"?}
The phenomenon of higher volatility of readings in the lower end of fluorescence measurements is well documented and comes from higher influence of small changes on the percentage error in that range.
Overall, 79\% and 27\% of our samples shows relative error of below 40\% and 20\% respectively which we find satisfactory for the purposes of this study.
